### 2018-07-01 paper/3-abstractions.tex
Switch to new track function

So much for applicative tasks. What about monadic tasks with dynamic
dependencies? As we have seen in~\S\ref{sec-background-shake}, dynamic
dependencies need to be tracked too. This cannot be done statically; notice that
we cannot apply the function \hs{dependencies} to a \hs{Task}~\hs{Monad} because
the \hs{Const} functor has no \hs{Monad} instance. We need to run a monadic task
on a store with concrete values, which will determine the discovered
dependencies. Accordingly, we introduce the function \hs{track}: a combination
of \hs{compute} and \hs{dependencies} that computes both the resulting value and
the list of its dependencies in an arbitrary monadic context~\hs{m}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Control.Monad.Writer
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Monad m => Task Monad k v -> (k -> m v) -> m (v, [k])
track task fetch = runWriterT $ run task trackingFetch
  where
    trackingFetch :: k -> WriterT [k] m v
    trackingFetch k = tell [k] >> lift (fetch k)
\end{minted}
\vspace{1mm}

\noindent
This implementation uses the standard Haskell \hs{WriterT} \emph{monad
transformer}~\cite{liang1995monad}, which allows us
to record additional information -- a list of keys of type \hs{[@@k]} -- when
computing a task in an arbitrary monad~\hs{m}. We substitute the given \hs{fetch}
with a \hs{trackingFetch} that, in addition to fetching a value, tracks the
corresponding key. The \hs{task} returns the value of type
\hs{WriterT}~\hs{[@@k]}~\hs{m}~\hs{v}, which we unwrap by \hs{runWriterT}. We
will use \hs{track} when implementing monadic build systems,
see~\S\ref{sec-implementation-shake}; here we show an example of \hs{track}ing
monadic tasks when \hs{m}~\hs{=}~\hs{IO}.

\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ fetchIO k = do putStr (k ++ ": "); read <$> getLine
@\ghci@ track (fromJust $ sprsh2 "B1") fetchIO
C1: 1
B2: 10
(10,["C1","B2"])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ track (fromJust $ sprsh2 "B1") fetchIO
C1: 2
A2: 20
(20,["C1","A2"])
\end{minted}
\vspace{1mm}

>>>>

So much for applicative tasks. What about monadic tasks with dynamic
dependencies? As we have seen in~\S\ref{sec-background-shake}, dynamic
dependencies need to be tracked too. This cannot be done statically; notice that
we cannot apply the function \hs{dependencies} to a \hs{Task}~\hs{Monad} because
the \hs{Const} functor has no \hs{Monad} instance. We need to run a monadic task
on a store with concrete values, which will determine the discovered
dependencies. Accordingly, we introduce the function \hs{track}: a combination
of \hs{compute} and \hs{dependencies} that computes both the resulting value and
the list of its dependencies (key/value pairs) in an arbitrary monadic
context~\hs{m}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Control.Monad.Writer
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Monad m => Task Monad k v -> (k -> m v) -> m (v, [(k, v)])
track task fetch = runWriterT $ run task trackingFetch
  where
    trackingFetch :: k -> WriterT [(k, v)] m v
    trackingFetch k = do v <- lift (fetch k); tell [(k, v)]; return v
\end{minted}
\vspace{1mm}

\noindent
This implementation uses the standard Haskell \hs{WriterT} \emph{monad
transformer}~\cite{liang1995monad}, which allows us to record additional
information -- a list of key/value pairs \hs{[(@@k,}~\hs{v)]} -- when computing
a task in an arbitrary monad~\hs{m}. We substitute the given \hs{fetch} with a
\hs{trackingFetch} that, in addition to fetching a value, tracks the
corresponding key/value pair. The \hs{task} returns the value of type
\hs{WriterT}~\hs{[(@@k,}~\hs{v)]}~\hs{m}~\hs{v}, which we unwrap by
\hs{runWriterT}. We will use \hs{track} when implementing monadic build systems
with dynamic dependencies, see~\S\ref{sec-implementation-shake}. Here we show an
example of \hs{track}ing monadic tasks when \hs{m}~\hs{=}~\hs{IO}.

\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ fetchIO k = do putStr (k ++ ": "); read <$> getLine
@\ghci@ track (fromJust $ sprsh2 "B1") fetchIO
C1: 1
B2: 10
(10,[("C1",1),("B2",10)])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ track (fromJust $ sprsh2 "B1") fetchIO
C1: 2
A2: 20
(20,[("C1",2),("A2",20)])
\end{minted}
\vspace{1mm}
